---
title: "Exercises 6"
author: "Shane Hauck"
date: "2/27/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Exercise 1. In the tennis app, change the histogram input from a sliderInput() to a numericInput().

```{r}
library(tidyverse)

atp_df <- read_csv("data/atp_matches_2019.csv")
wta_df <- read_csv("data/wta_matches_2019.csv")
both_df <- bind_rows(atp_df, wta_df)

both_long <- both_df %>% pivot_longer(c(winner_name, loser_name))

both_n50 <- both_long %>% group_by(value) %>% count() %>%
  filter(n > 50)

major_tennis <- semi_join(both_long, both_n50, by = c("value"))
major_tennis <- major_tennis %>% mutate(w_svperc = 100 * w_1stIn / w_svpt,
                                        l_svperc = 100 * l_1stIn / l_svpt,
                                        w_firstwon = 100 * w_1stWon / w_1stIn,
                                        l_firstwon = 100 * l_1stWon / l_1stIn,
                                        w_secondwon = 100 * w_2ndWon / (w_svpt - w_1stIn),
                                        l_secondwon = 100 * l_2ndWon / (l_svpt - l_1stIn))


major_tennis_w <- major_tennis %>% filter(name == "winner_name")
major_tennis_l <- major_tennis %>% filter(name == "loser_name")

w_small <- major_tennis_w %>% select(value, winner_seed, w_ace, w_df, w_svperc,
                                     w_firstwon, w_secondwon) %>%
  rename(seed = winner_seed, ace = w_ace, df = w_df, svperc = w_svperc,
         firstwon = w_firstwon, secondwon = w_secondwon)

l_small <- major_tennis_l %>% select(value, loser_seed, l_ace, l_df, l_svperc, l_firstwon, l_secondwon)  %>%
  rename(seed = loser_seed, ace = l_ace, df = l_df, svperc = l_svperc,
         firstwon = l_firstwon, secondwon = l_secondwon)

df <- bind_rows(w_small, l_small) %>%
  rename(player = "value")
df

ggplot(data = df %>% filter(player == "Roger Federer"),
       aes(x = ace)) +
  geom_histogram()




library(shiny)
var_choices <- names(df)[3:7]

ui <- fluidPage(
  sidebarLayout(sidebarPanel(
    selectizeInput(inputId = "playerchoice", 
                   label = "Choose a player", 
                   choices = levels(factor(df$player)),
                   selected = "Roger Federer"),
    radioButtons("varchoice", label = "Choose a stat",
                 choices = var_choices)),
    numericInput("binnumber", label = "Choose a Number of Bins", 
                min = 1, max = 50, value = 15)),
  mainPanel(plotOutput("histgraph"))
)

server <- function(input, output, session) {
  df_sub <- reactive({
    df %>% filter(player == input$playerchoice)
  })
  hist_plot <- reactive({
    # ggplot(df_sub(), aes_string(x = input$varchoice)) +
    # geom_histogram(colour = "black", fill = "white", bins = 15)
    ggplot(df_sub(), aes(x = .data[[input$varchoice]])) +
      geom_histogram(colour = "black", fill = "white", bins = input$binnumber) +
      theme_grey(base_size = 22)
  })
  
  output$histgraph <- renderPlot({
    hist_plot()
  }) 
}

shinyApp(ui, server)
```


Exercise 2. In the tennis app, change the histogram input back to sliderinput(). Then, apply 2.2.8 Exercise 3 in Mastering Shiny to the tennis app slider.

```{r}


ui <- fluidPage(
  sidebarLayout(sidebarPanel(
    selectizeInput(inputId = "playerchoice", 
                   label = "Choose a player", 
                   choices = levels(factor(df$player)),
                   selected = "Roger Federer"),
    radioButtons("varchoice", label = "Choose a stat",
                 choices = var_choices)),
    sliderInput("binnumber", label = "Choose a Number of Bins", 
                min = 0, max = 100, value = 0, step = 5, animate = TRUE)),
  mainPanel(plotOutput("histgraph"))
  )

server <- function(input, output, session) {
  df_sub <- reactive({
    df %>% filter(player == input$playerchoice)
  })
  hist_plot <- reactive({
    # ggplot(df_sub(), aes_string(x = input$varchoice)) +
    # geom_histogram(colour = "black", fill = "white", bins = 15)
    ggplot(df_sub(), aes(x = .data[[input$varchoice]])) +
      geom_histogram(colour = "black", fill = "white", bins = input$binnumber) +
      theme_grey(base_size = 22)
  })
  
  output$histgraph <- renderPlot({
    hist_plot()
  }) 
}

shinyApp(ui, server)

```

Exercise 3. In the SLU majors app, change the major input to radioButtons() so that the user can only select mathematics, statistics, or computer science as the major.

```{r}
library(tidyverse)
library(readxl)
df <- read_excel("data/slu_graduates_17_21.xlsx")

## fixes error in the data
df <- df %>% mutate(across(everything(),
                           .fns = ~replace(., . ==  "STATS" , "STAT")))

df_long <- df %>% pivot_longer(3:8, names_to = "type", values_to = "discipline")
df_major <- df_long %>% 
  filter(type == "major1" | type == "major2" | type == "major3") %>%
   filter(discipline == "STAT" | discipline == "CS" | discipline == "MATH")

df_stat <- df_major %>% filter(discipline == "STAT") 
df_statfull <- semi_join(df_long, df_stat, by = "adm_id") %>%
  filter(type == "major1" |
           type == "major2" | 
           type == "major3")
 

df_nostat <- df_statfull %>% filter(discipline != "STAT" &
                                      !is.na(discipline)) %>%
  group_by(discipline) %>%
  summarise(nstudent = n()) %>%
  mutate(discipline = fct_reorder(discipline, nstudent))
ggplot(data = df_nostat, aes(x = discipline, y = nstudent)) +
  geom_col() +
  coord_flip()


## step 2: input
library(shiny)

df_nomiss <- df_major %>% filter(!is.na(discipline)) %>%
  mutate(discipline = factor(discipline))
majors <- levels(df_nomiss$discipline)

ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(radioButtons(inputId = "majorchoice", label = "Choose a Major",
                             choices = factor(majors))),
    mainPanel(plotOutput("majorplot"))
  )
)

server <- function(input, output, session) {
  major_reactive_df <- reactive({
    
  df_selected <- df_major %>% filter(discipline == input$majorchoice) 
  df_selectedfull <- semi_join(df_long, df_selected, by = "adm_id") %>%
    filter(type == "major1" |
             type == "major2" | 
             type == "major3")
  
  df_noselected <- df_selectedfull %>% filter(discipline != input$majorchoice &
                                        !is.na(discipline)) %>%
    group_by(discipline) %>%
    summarise(nstudent = n()) %>%
    mutate(discipline = fct_reorder(discipline, nstudent))})
  
  major_plot <- reactive({
    ggplot(data = major_reactive_df(), aes(x = discipline, y = nstudent)) +
      geom_col() +
      coord_flip()
  })
  
  output$majorplot <- renderPlot({
    major_plot()
  })
}

shinyApp(ui, server)

```

Exercise 4. In the SLU majors app, add a table output below the plot that shows the number of Female majors and the number of Male majors for a major that the user selects.

```{r}
library(tidyverse)
library(readxl)
df <- read_excel("data/slu_graduates_17_21.xlsx")

## fixes error in the data
df <- df %>% mutate(across(everything(),
                           .fns = ~replace(., . ==  "STATS" , "STAT")))

df_long <- df %>% pivot_longer(3:8, names_to = "type", values_to = "discipline")
df_major <- df_long %>% 
  filter(type == "major1" | type == "major2" | type == "major3")

df_stat <- df_major %>% filter(discipline == "STAT") 
df_statfull <- semi_join(df_long, df_stat, by = "adm_id") %>%
  filter(type == "major1" |
           type == "major2" | 
           type == "major3")

df_nostat <- df_statfull %>% filter(discipline != "STAT" &
                                      !is.na(discipline)) %>%
  group_by(discipline) %>%
  summarise(nstudent = n()) %>%
  mutate(discipline = fct_reorder(discipline, nstudent))
ggplot(data = df_nostat, aes(x = discipline, y = nstudent)) +
  geom_col() +
  coord_flip()


## step 2: input
library(shiny)

df_nomiss <- df_major %>% filter(!is.na(discipline)) %>%
  mutate(discipline = factor(discipline))
majors <- levels(df_nomiss$discipline)

ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(selectInput(inputId = "majorchoice", label = "Choose a Major",
                             choices = factor(majors))),
    mainPanel(plotOutput("majorplot"), dataTableOutput("gendertable"))
  )
)

server <- function(input, output, session) {
  major_reactive_df <- reactive({
    
  df_selected <- df_major %>% filter(discipline == input$majorchoice) 
  df_selectedfull <- semi_join(df_long, df_selected, by = "adm_id") %>%
    filter(type == "major1" |
             type == "major2" | 
             type == "major3")
  
  df_noselected <- df_selectedfull %>% filter(discipline != input$majorchoice &
                                        !is.na(discipline)) %>%
    group_by(discipline) %>%
    summarise(nstudent = n()) %>%
    mutate(discipline = fct_reorder(discipline, nstudent))
  })
  

  
  major_plot <- reactive({
    ggplot(data = major_reactive_df(), aes(x = discipline, y = nstudent)) +
      geom_col() +
      coord_flip()
  })
  
  gender_table <- reactive({
     df_selected <- df_major %>% filter(discipline == input$majorchoice) 
    df_selectedfull <- semi_join(df_long, df_selected, by = "adm_id") %>%
      filter(type == "major1" |
             type == "major2" | 
             type == "major3")
  
    df_noselected <- df_selectedfull %>% filter(discipline != input$majorchoice &
                                        !is.na(discipline)) %>%
      group_by(sex) %>%
      summarise(nstudent = n())
  })
  
  output$majorplot <- renderPlot({
    major_plot()
  })
  
  output$gendertable <- renderDataTable({
    gender_table()
  })
}

shinyApp(ui, server)
```  

 