---
title: "04_Quiz_TakeHome"
author: "Shane Hauck"
date: "2/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

“All work presented is my own, and I have followed all rules for collaboration.”
- Shane Hauck


Use the leaflet package and the us.cities data set from the maps package to make a Zoom-able map of the 50 U.S. capital cities. When you click on a marker for a capital city, both the city name and state as well as the pop (population) should be displayed.

Note that, in the us.cities data set, state capitals are given a value of 2 in the capital variable.

A second note is that if you cannot figure out how to get two “things” (in this case name and population) to display in Leaflet, you would just lose a single point for the quiz (so don’t spend too much time on that part unless you are inclined to do so).

```{r}
library(tidyverse)
library(maps)
library(leaflet)
```


```{r}
capitals_df <- maps::"us.cities" %>% filter(capital == 2)
```

```{r}
content <- capitals_df %>%
  mutate(popup = paste("Capital:", capitals_df$name, "<br>",
                       "Population:", capitals_df$pop, "<br>"))
```

```{r}
capital_map <- leaflet(capitals_df) %>%
  setView(lng = -98.583, lat = 39.833, zoom = 3) %>%
  addTiles() %>%
  addProviderTiles(providers$Wikimedia) %>%
  addMarkers(lng = capitals_df$long, lat = capitals_df$lat,
             clusterOptions = markerClusterOptions(),
             popup = content$popup)
capital_map
```
